I'd first like to thank my friends and fellow Flight Lab people. Each of you brought something good to my time at Poly. B$^2$ and Cory, you guys were awesome roommates and made the Stafford house fun. Alex, Brian, Christian, Trevor, Mike, and the rest of the DBF/Flight Lab crew, you guys made that window-less lab bearable, and occasionally downright enjoyable. \#W9W

I'd also like to thank my committee. Throughout my Cal Poly career every member of the committee has helped me immensely. Dr. Colvin, hosting Akaflieg barbecues at your hangar introduced my to a side of aviation I hadn't seen, and gave me passion for aircraft and flight. Dr. Mehiel, you've provided countless guidance throughout school, ranging from classes and scheduling to Lyapunov stability and Kalman filters. Dr. Westphal, the Edward's project was a great experience for me, and you've been incredibly supportive with anything and everything I've needed help with since. And Dr. McDonald, you've given me countless opportunities, including DBF, Edwards, LENR, Senior Design, R/MAX, this thesis, ... For five and a half years, you've guided me through challenging problems, taught me to think like an engineer, and ignored DBF t-shirts and team names. Thank you, it's been a pleasure.

Finally, my family. The entire family (aunts, uncles, grandparents, etc.) have supported me the whole way through, and I'm very grateful for that. I can't put into words the opportunities and support my parents and sister have given me. Mom, Dad, Jackie: Thank you. You guys are awesome.