\chapter{Flight Test Procedure}
\label{sect:flightTestDocs}
This section documents proper flight test procedure, which has been learned from extensive flight testing experience and many crashed vehicles. It is split into three main time periods:
\begin{enumerate}
\item Pre-Flight Preparation
\item Flying Field Procedure
\item Post-Flight 
\end{enumerate}
The testing procedure is split into these time categories to ensure the testing is efficient and that any unforeseen circumstances can be dealt with quickly. Specifically , this guide applies to the Cal Poly Flight Lab testing a vehicle at the Cal Poly Educational Flight Range.

\section{Pre-Flight Preparation}
The preparation work required for a test flight is often overlooked, and this section will document how to effectively prepare to flight test a vehicle. The main purpose of pre-flight preparation is to minimize the possibility of problems that might force a test to be canceled after already going to the field.
\subsection{Day Before Test}
The days before a flight test are critical to ensuring the test is completed successfully. The following should be done at least one day before the the test is scheduled:
\begin{enumerate}
\item Charge all flight battery packs, including any receiver or auxiliary packs.
\item Charge the transmitter battery.
\item Ensure the airframe is structurally sound (wing tip test minimum.)
\item Verify radio system communicates properly, and the correct fail-safe is in place. \textbf{Important}: If the receiver has been used by Design/Build/Fly, the fail-safe must be changed to normal mode.
\item Verify control surface deflections matches desired directions, and all radio mixes work.
\item Verify the motor/propeller spin in the correct direction.
\item Pack a flight box, containing any necessary tools (recommend: screw drivers; masking, painter's, and strapping tape; razor blades; CA glue and kicker; spare propellers; paper and pencil; as a minimum.)
\item Check the weather. The closest monitoring station to the field is \href{http://www.wunderground.com/cgi-bin/findweather/getForecast?query=35.326\%2C-120.738\&sp=KCASANLU17}{KCASANLU17}.It is generally better to flight test early in the morning, since there will be less people at the field, winds will be calmer, and the sun won't be as harsh in the pilot's eyes.
\item Check the SLO Flyers' flight schedule. Some days are reserved for certain events (glider competitions, etc.) and these need to be worked around.
\item Make sure all required personnel know when and where to meet, and there is sufficient transportation to get to the field.
\item Create flight documentation that clearly lays out the test goals and how they will be accomplished. Print copies for all personnel.
\end{enumerate}
\subsection{Day Of Test}
The morning of the flight test, the person responsible for the test should arrive early enough to accomplish the following, before leaving campus:
\begin{enumerate}
\item Pack flight batteries into flight box, including receiver and auxiliary packs.
\item Pack battery charging equipment, with adapters and leads, if necessary.
\item Pack transmitter into box.
\item Double check control surface deflections and radio link.
\item Do a full system check, potentially including a short taxi test in the quad.
\item Do a final check that all equipment made it into vehicles, before leaving the lab.
\end{enumerate}

\section{Flying Field Procedure}
With the pre-flight preparation completed, the testing at the flying field should be fairly event free. Any problems that occur should be either fixable with the minimum supplies in the flight box, or the test should be canceled to minimize risk, and repairs done at the lab. Specific procedures at the flight field will depend on the test being conducted, but below are steps that apply to nearly all tests before flight.
\begin{enumerate}
\item Verify structural integrity using a wing tip test.
\item Verify motor/propeller are spinning in the correct direction.
\item Verify control surface deflections match desired directions.
\item Verify radio link and fail safe mode by completing a range check.
\item Verify center of gravity is at an appropriate location.
\item Create flight timer on the transmitter so the pilot knows how long the aircraft has been flying.
\item Document any required data before the flight, such as aircraft weight and geometry.
\item Document current weather conditions (wind speed and direction, temperature, humidity, barometric pressure) using the Kestrel portable weather station.
\end{enumerate}

\section{Post-Flight}
Upon successful completion of a flight test, all data should be saved to a computer. The flight test cards, pre-flight data documentation, and any notes should be scanned and saved with the flight data. This data should be archived in a .zip file, with the date attached.

If the test flight resulted in a crash, any available data should still be saved. Any video or pictures of the flight should be saved as well, to aid in isolating what caused the crash. If useful, pictures should be taken of the crash site. Afterwards, all pieces of the aircraft should be returned to the lab, where the root cause of the accident should be determined.

After the root cause has been determined, remove all electronics from the vehicle, including batteries, speed controller, motor, servos, and receiver. If the vehicle was a complete loss or went into water, through these components away: \textbf{do not} return to lab supplies. If the vehicle was not a complete loss, verify all components work properly.